%%
%% This is file `apsstempl.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% stepe.dtx  (with options: `apsstempl')
%% 
%%     This work has been partially funded by the US government
%%  and is not subject to copyright.
%% 
%%     This program is provided under the terms of the
%%  LaTeX Project Public License distributed from CTAN
%%  archives in directory macros/latex/base/lppl.txt.
%% 
%%  Author: Peter Wilson (CUA and NIST)
%%          now at: peter.r.wilson@boeing.com
%% 
\ProvidesFile{apsstempl.tex}[2001/07/16 AP SUP/SUB templates boilerplate]
\typeout{apsstempl.tex [2001/07/16 AP SUP/SUB templates boilerplate]}

\sssclause{SUBTYPE}

    The SUBTYPE mapping template specifies a reference to the mapping of
a subtype of the current application object. Several such references may
be included for one supertype application object.

\begin{anote} This template definition only consists of a template signature,
    there is no matching template body. The template is included to ease the
    automatic processing of the mapping specification.
\end{anote}

\signature

/SUBTYPE(application\_object)/

\parameters

application\_object: the application object that is a subtype of the current
                     supertype application object and that has the entire
                     or a part of the mapping specification of this
                     supertype.

\sssclause{SUPERTYPE}

    The SUPERTYPE mapping template specifies a reference to the mapping of
a supertype of the current application object. Several such references may
be included for the subtype application object.

\begin{anote} This template only consists of a signature,
    there is no matching body. The template is included to ease the
    automatic processing of the mapping specification.
\end{anote}

\signature

/SUPERTYPE(application\_object)/

\parameters

application\_object: the application object that is a supertype of the current
                     subtype application object and that has the entire
                     or a part of the mapping specification of this
                     subtype.

\endinput
%%
%% End of file `apsstempl.tex'.
