% ======================================================================
% common-15.tex
% Copyright (c) Markus Kohm, 2001-2012
%
% This file is part of the LaTeX2e KOMA-Script bundle.
%
% This work may be distributed and/or modified under the conditions of
% the LaTeX Project Public License, version 1.3c of the license.
% The latest version of this license is in
%   http://www.latex-project.org/lppl.txt
% and version 1.3c or later is part of all distributions of LaTeX 
% version 2005/12/01 or later and of this work.
%
% This work has the LPPL maintenance status "author-maintained".
%
% The Current Maintainer and author of this work is Markus Kohm.
%
% This work consists of all files listed in manifest.txt.
% ----------------------------------------------------------------------
% common-15.tex
% Copyright (c) Markus Kohm, 2001-2012
%
% Dieses Werk darf nach den Bedingungen der LaTeX Project Public Lizenz,
% Version 1.3c, verteilt und/oder veraendert werden.
% Die neuste Version dieser Lizenz ist
%   http://www.latex-project.org/lppl.txt
% und Version 1.3c ist Teil aller Verteilungen von LaTeX
% Version 2005/12/01 oder spaeter und dieses Werks.
%
% Dieses Werk hat den LPPL-Verwaltungs-Status "author-maintained"
% (allein durch den Autor verwaltet).
%
% Der Aktuelle Verwalter und Autor dieses Werkes ist Markus Kohm.
% 
% Dieses Werk besteht aus den in manifest.txt aufgefuehrten Dateien.
% ======================================================================
%
% Paragraphs that are common for several chapters of the KOMA-Script guide
% Maintained by Markus Kohm
%
% ----------------------------------------------------------------------
%
% Absaetze, die mehreren Kapiteln der KOMA-Script-Anleitung gemeinsam sind
% Verwaltet von Markus Kohm
%
% ======================================================================

\ProvidesFile{common-15.tex}[2012/05/15 KOMA-Script guide (common paragraphs)]
\translator{Gernot Hassenpflug\and Markus Kohm\and Krickette Murabayashi}

% Date of the translated German file: 2011/09/21

\makeatletter
\@ifundefined{ifCommonmaincls}{\newif\ifCommonmaincls}{}%
\@ifundefined{ifCommonscrextend}{\newif\ifCommonscrextend}{}%
\@ifundefined{ifCommonscrlttr}{\newif\ifCommonscrlttr}{}%
\@ifundefined{ifIgnoreThis}{\newif\ifIgnoreThis}{}%
\makeatother


\section{Dicta}
\label{sec:\csname label@base\endcsname.dictum}%
\ifshortversion\IgnoreThisfalse\IfNotCommon{maincls}{\IgnoreThistrue}\fi%
\ifIgnoreThis %+++++++++++++++++++++++++++++++++++++++++++++ nicht maincls +
It applies, mutatis mutandis, what is described in
\autoref{sec:maincls.dictum}.
\IfCommon{scrextend}{%
  However, \Package{scrextend} does not support the commands
  \Macro{setchapterpreamble} and \Macro{setpartpreamble}. You should read
  the manual of the used class, if you want to know, if that class does
  support similar commands.}
\else %------------------------------------------------------- nur maincls -
\BeginIndex{}{smart slogan}%
\BeginIndex{}{dictum}%
\BeginIndex{}{excerpt}%

Sometimes you may find a dictum, a kind of smart slogan or excerpt, often ragged
left above or below the heading of a chapter or section. The text and the
source of the slogan often use special styles.


\begin{Declaration}
  \Macro{dictum}\OParameter{author}\Parameter{dictum}\\
  \Macro{dictumwidth}\\
  \Macro{dictumauthorformat}\Parameter{author}\\
  \Macro{dictumrule}\\
  \Macro{raggeddictum}\\
  \Macro{raggeddictumtext}\\
  \Macro{raggeddictumauthor}
\end{Declaration}%
\BeginIndex{Cmd}{dictum}%
\BeginIndex{Cmd}{dictumwidth}%
\BeginIndex{Cmd}{dictumauthorformat}%
\BeginIndex{Cmd}{dictumrule}%
\BeginIndex{Cmd}{raggeddictum}%
\BeginIndex{Cmd}{raggeddictumtext}%
\BeginIndex{Cmd}{raggeddictumauthor}%
The command \Macro{dictum} inserts such a dictum.
\IfCommon{maincls}{This\textnote{Hint!} macro can be used as obligatory
  argument of either the command \Macro{setchapterpreamble} or
  \Macro{setpartpreamble}. However, this is not obligatory.\par}%
The dictum together with an optional \PName{author} is inserted in a
\Macro{parbox}\IndexCmd{parbox} (see \cite{latex:usrguide}) of width
\Macro{dictumwidth}. Yet \Macro{dictumwidth} is not a length which can be set
with \Macro{setlength}. It is a macro that can be redefined using
\Macro{renewcommand}. Default setting is \verb;0.3333\textwidth;, which is a
third of the textwidth. The box itself is positioned with the command
\Macro{raggeddictum}. Default here is \Macro{raggedleft}\IndexCmd{raggedleft},
that is, right justified.  The command \Macro{raggeddictum} can be redefined
using \Macro{renewcommand}.

Within the box the \PName{dictum} is set using \Macro{raggeddictumtext}.
Default setting is \Macro{raggedright}\IndexCmd{raggedright}, that is, left
justified. Similarly to \Macro{raggeddictum} this can be redefined with
\Macro{renewcommand}.  The output uses the default font setting for the
element \FontElement{dictumtext}, which can be changed with the commands
\Macro{setkomafont} and \Macro{addtokomafont} (see \autoref{sec:\csname
  label@base\endcsname.textmarkup}, \autopageref{desc:\csname
  label@base\endcsname.cmd.setkomafont}). Default settings are listed in
\autoref{tab:maincls.dictumfont}%
\IfNotCommon{maincls}{, \autopageref{tab:maincls.dictumfont}}%
.

\ifCommonmaincls
\begin{table}
%  \centering%
  \KOMAoptions{captions=topbeside}%
  \setcapindent{0pt}%
%  \caption
  \begin{captionbeside}{Default settings for the elements of a dictum}
    [l]
  \begin{tabular}[t]{ll}
    \toprule
    Element & Default \\
    \midrule
    \FontElement{dictumtext} &
    \Macro{normalfont}\Macro{normalcolor}\Macro{sffamily}\Macro{small}\\
    \FontElement{dictumauthor} &
    \Macro{itshape}\\
    \bottomrule
  \end{tabular}
  \end{captionbeside}
  \label{tab:maincls.dictumfont}
\end{table}
\fi

If there is an \PName{author} name, it is separated from the \PName{dictum} by
a rule to the full width of the
\Macro{parbox}. This\IfCommon{maincls}{\ChangedAt{v3.10}{\Class{scrbook}\and
    \Class{scrreprt}\and \Class{scrartcl}}}%
\IfCommon{scrextend}{\ChangedAt{v3.10}{\Package{scrextend}}} rule is defined
as vertical object to command \Macro{dictumrule}:
% Satzkorrektur listings
\begin{lstcode}[belowskip=\dp\strutbox]
  \newcommand*{\dictumrule}{\vskip-1ex\hrulefill\par}
\end{lstcode}

The alignment is defined with \Macro{raggeddictumauthor}. Default is
\Macro{raggedleft}. This command can also be redefined using
\Macro{renewcommand}. The format of the output is defined with
\Macro{dictumauthorformat}. This macro expects the \Macro{author} as
argument. As default \Macro{dictumauthorformat} is defined as:
% Satzkorrektur listings
\begin{lstcode}[belowskip=\dp\strutbox]
  \newcommand*{\dictumauthorformat}[1]{(#1)}
\end{lstcode}
Thus the \PName{author} is set enclosed in rounded parentheses. For
the element \FontElement{dictumauthor}, a different font than for the
element \FontElement{dictumtext} can be defined. Default settings are
listed in \autoref{tab:maincls.dictumfont}. Changes can be made using
the commands \Macro{setkomafont} and \Macro{addtokomafont} (see
\autoref{sec:\csname label@base\endcsname.textmarkup},
\autopageref{desc:\csname label@base\endcsname.cmd.setkomafont}).%
%
\ifCommonmaincls

If \Macro{dictum} is used within the macro \Macro{setchapterpreamble}
or \Macro{setpartpreamble} you have to take care of the following: the
horizontal positioning is always done with \Macro{raggeddictum}.
Therefore, the optional argument for horizontal positioning which is
implemented for these two commands has no effect. \Macro{textwidth} is
not the width of the whole text corpus but the actually used text
width.  If \Macro{dictumwidth} is set to \verb;.5\textwidth; and
\Macro{setchapterpreamble} has an optional width of
\verb;.5\textwidth; too, you will get a box with a width one quarter
of the text width.  Therefore, if you use \Macro{dictum} it is
recommended to refrain from setting the optional width for
\Macro{setchapterpreamble} or \Macro{setpartpreamble}.

If\textnote{Hint!} you have more than one dictum, one under another, you should
separate them by an additional vertical space, easily accomplished using the
command \Macro{bigskip}\IndexCmd{bigskip}.

\begin{Example}
  You are writing a chapter on an aspect of weather forecasting. You
  have come across an aphorism which you would like to place at the
  beginning of the chapter beneath the heading. You could write:
\begin{lstcode}
  \setchapterpreamble[u]{%
    \dictum[Anonymous]{Forecasting is the art of saying
            what is going to happen and then to explain
            why it didn't.}}
  \chapter{Weather forecasting}
\end{lstcode}
  The output would look as follows:
  \begin{ShowOutput}
    {\usekomafont{disposition}\usekomafont{chapter}%
      17\enskip Weather forecasting\par} \vspace{\baselineskip}
    \dictum[Anonymous]{Forecasting is the art of saying what is going to
    happen and then to explain why it didn't.}
  \end{ShowOutput}

  If you would rather the dictum span only a quarter of the text width
  rather than one third you can redefine \Macro{dictumwidth}:
\begin{lstcode}
  \renewcommand*{\dictumwidth}{.25\textwidth}
\end{lstcode}
\end{Example}

For a somewhat more sophisticated formatting of left- or right-aligned
paragraphs including hyphenation you can use the
package~\Package{ragged2e}~\cite{package:ragged2e}.%
\fi
%
\EndIndex{Cmd}{dictum}%
\EndIndex{Cmd}{dictumwidth}%
\EndIndex{Cmd}{dictumauthorformat}%
\EndIndex{Cmd}{dictumrule}%
\EndIndex{Cmd}{raggeddictum}%
\EndIndex{Cmd}{raggeddictumtext}%
\EndIndex{Cmd}{raggeddictumauthor}%
%
\EndIndex{}{excerpt}%
\EndIndex{}{smart slogan}%
\EndIndex{}{dictum}%
\fi  %**************************************************** Ende nur maincls *


%%% Local Variables:
%%% mode: latex
%%% coding: us-ascii
%%% TeX-master: "../guide"
%%% End:
