% ======================================================================
% common-7.tex
% Copyright (c) Markus Kohm, 2001-2012
%
% This file is part of the LaTeX2e KOMA-Script bundle.
%
% This work may be distributed and/or modified under the conditions of
% the LaTeX Project Public License, version 1.3c of the license.
% The latest version of this license is in
%   http://www.latex-project.org/lppl.txt
% and version 1.3c or later is part of all distributions of LaTeX 
% version 2005/12/01 or later and of this work.
%
% This work has the LPPL maintenance status "author-maintained".
%
% The Current Maintainer and author of this work is Markus Kohm.
%
% This work consists of all files listed in manifest.txt.
% ----------------------------------------------------------------------
% common-7.tex
% Copyright (c) Markus Kohm, 2001-2012
%
% Dieses Werk darf nach den Bedingungen der LaTeX Project Public Lizenz,
% Version 1.3c, verteilt und/oder veraendert werden.
% Die neuste Version dieser Lizenz ist
%   http://www.latex-project.org/lppl.txt
% und Version 1.3c ist Teil aller Verteilungen von LaTeX
% Version 2005/12/01 oder spaeter und dieses Werks.
%
% Dieses Werk hat den LPPL-Verwaltungs-Status "author-maintained"
% (allein durch den Autor verwaltet).
%
% Der Aktuelle Verwalter und Autor dieses Werkes ist Markus Kohm.
% 
% Dieses Werk besteht aus den in manifest.txt aufgefuehrten Dateien.
% ======================================================================
%
% Paragraphs that are common for several chapters of the KOMA-Script guide
% Maintained by Markus Kohm
%
% ----------------------------------------------------------------------
%
% Absaetze, die mehreren Kapiteln der KOMA-Script-Anleitung gemeinsam sind
% Verwaltet von Markus Kohm
%
% ======================================================================

\ProvidesFile{common-7.tex}[2012/05/15 KOMA-Script guide (common paragraphs)]
\translator{Markus Kohm\and Krickette Murabayashi}

% Date of translated German file: 2011/08/26

\makeatletter
\@ifundefined{ifCommonmaincls}{\newif\ifCommonmaincls}{}%
\@ifundefined{ifCommonscrextend}{\newif\ifCommonscrextend}{}%
\@ifundefined{ifCommonscrlttr}{\newif\ifCommonscrlttr}{}%
\@ifundefined{ifIgnoreThis}{\newif\ifIgnoreThis}{}%
\makeatother


\section{Detection of Odd and Even Pages}
\label{sec:\csname label@base\endcsname.oddOrEven}%
\ifshortversion\IgnoreThisfalse\IfNotCommon{maincls}{\IgnoreThistrue}\fi%
\ifIgnoreThis %+++++++++++++++++++++++++++++++++++++++++++++ nicht maincls +
What is described in
\autoref{sec:maincls.oddOrEven} applies, mutatis mutandis.
\else %------------------------------------------------------- nur maincls -
\BeginIndex{}{page>odd}%
\BeginIndex{}{page>even}%

In double-sided documents we distinguish left and right pages. Left pages
always have an even page number, right pages always have an odd page
number. % For this they are mostly called even and odd pages so they
        % are
Thus, they are most often referred to as even and odd pages
in this
\iffree{guide}{book}. %
\IfCommon{maincls}{This also means that the detection
  of a left or right page is same as detection of even and odd page
  numbers.

% Umbruchkorrekturtext
  \iftrue%
  There's no distinction in left and right pages in single-sided
  documents. Nevertheless there are pages with even or odd page numbers.%
  \fi}%
\IfCommon{scrlttr2}{Letters will be set single-side mostly. Nevertheless,
  printing letters with single-side layout using both paper sides or
  exceptionally generating real double-side letters it may be useful to know
  whether producing currently an even or odd page.}%

\begin{Declaration}
  \Macro{ifthispageodd}\Parameter{true part}\Parameter{false part}
\end{Declaration}%
\BeginIndex{Cmd}{ifthispageodd}%
If one wants to find out with {\KOMAScript} whether a text falls on an
even or odd page, one can use the \Macro {ifthispageodd} command. The
\PName{true part} argument is executed only if the command falls on an odd
page.  Otherwise the \PName{false part} argument is executed.

\begin{Example}
  Assume that you simply want to show whether a text will be placed onto an
  even or odd page. You may achieve that using
\begin{lstcode}[belowskip=-\baselineskip]
  This page has an \ifthispageodd{odd}{even} 
  page number.
\end{lstcode}
  which will result in the output
  \begin{quote}
    This page has an \ifthispageodd{odd}{even} page number.
  \end{quote}
\end{Example}

Because the \Macro{ifthispageodd} command uses a mechanism that is
very similar to a label and a reference to it, at least two {\LaTeX}
runs are required after every text modification. Only then the
decision is correct. In the first run a heuristic is used to make the
first choice.

At \autoref{sec:maincls-experts.addInfos},
\autopageref{desc:maincls-experts.cmd.ifthispageodd} experts may find more
information about the problems detecting left and right pages or even and odd
page number.%
%
\EndIndex{Cmd}{ifthispageodd}%
%
\EndIndex{}{page>odd}%
\EndIndex{}{page>even}
\fi %**************************************************** Ende nur maincls *


%%% Local Variables:
%%% mode: latex
%%% coding: us-ascii
%%% TeX-master: "../guide"
%%% End:
