%
% Herein I define some safety advices which I had to include with my
% lab protocols in organic chemistry.
% The latest version will be made available at
% http://projects.thiemo.net/RundS/
% and distributed via CTAN
%
%
% Copyright (c) 1997-2005 Thiemo Nordenholz <nz@thiemo.net>
% All rights reserved.
% $Id: eng_rs.tex,v 1.6 2005/12/01 12:21:21 thiemo Exp $
%
% Redistribution and use in source and binary forms, with or without
% modification, are permitted provided that the following conditions
% are met:
% 1. Redistributions of source code must retain the above copyright
%    notice, this list of conditions and the following disclaimer.
% 2. Redistributions in binary form must reproduce the above copyright
%    notice, this list of conditions and the following disclaimer in the
%    documentation and/or other materials provided with the distribution.
%
% THIS SOFTWARE IS PROVIDED BY THE AUTHOR AND CONTRIBUTORS ``AS IS'' AND
% ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
% IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
% ARE DISCLAIMED.  IN NO EVENT SHALL THE AUTHOR OR CONTRIBUTORS BE LIABLE
% FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL
% DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS
% OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
% HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
% LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY
% OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF
% SUCH DAMAGE.
%
%

\message{Package 'eng_rs', 01.12.2005 (v1.3i), nz@thiemo.net}

% numbers are coded o, e, z, d, v, f, s, i, a, n
% combinations are prefixed with an extra-'c',
% for information and examples see README.

% define some 'R'-phrases:
% R1-9
\def\cre{R1 Explosive when dry.}
\def\crz{R2 Risk of explosion by shock, friction, fire or other 
sources of ignition.}
\def\crd{R3 Extreme risk of explosion by shock, friction, fire or
other sources of ignition.}
\def\crv{R4 Forms very sensitive explosive metallic compounds.}
\def\crf{R5 Heating may cause an explosion.}
\def\crs{R6 Explosive with or without contact with air.}
\def\cri{R7 May cause a fire.}
\def\cra{R8 Contact with combustible material may cause fire.}
\def\crn{R9 Explosive when mixed with combustible material.}
% R10-19
\def\creo{R10 Flammable.}
\def\cree{R11 Highly flammable.}
\def\crez{R12 Extremely flammable.}
\def\cred{R13 Extremely flammable liquefied gas.}
\def\crev{R14 Reacts violently with water.}
\def\cref{R15 Contact with water liberates highly flammable gases.}
\def\cres{R16 Explosive when mixed with oxidizing substances.}
\def\crei{R17 Spontaneously flammable in air.}
\def\crea{R18 In use, may form flammable/explosive vapour-air
mixture.}
\def\cren{R19 May form explosive peroxides.}
% R20-29
\def\crzo{R20 Harmful by inhalation.}
\def\crze{R21 Harmful in contact with skin.}
\def\crzz{R22 Harmful if swallowed.}
\def\crzd{R23 Toxic by inhalation.}
\def\crzv{R24 Toxic in contact with skin.}
\def\crzf{R25 Toxic if swallowed.}
\def\crzs{R26 Very toxic by inhalation.}
\def\crzi{R27 Very toxic in contact with skin.}
\def\crza{R28 Very toxic if swallowed.}
\def\crzn{R29 Contact with water liberates toxic gas.}
% R30-39
\def\crdo{R30 Can become highly flammable in use.}
\def\crde{R31 Contact with acids liberates toxic gas.}
\def\crdz{R32 Contact with acids liberates very toxic gas.}
\def\crdd{R33 Danger of cumulative effects.}
\def\crdv{R34 Causes burns.}
\def\crdf{R35 Causes severe burns.}
\def\crds{R36 Irritating to eyes.}
\def\crdi{R37 Irritating to respiratory system.}
\def\crda{R38 Irritating to skin.}
\def\crdn{R39 Danger of very serious irreversible effects.}
% R40-48
\def\crvo{R40 Possible risks of irreversible effects.}
\def\crve{R41 Risk of serious damage to eyes.}
\def\crvz{R42 May cause sensitization by inhalation.}
\def\crvd{R43 May cause sensitization by skin contact.}
\def\crvv{R44 Risk of explosion if heated under confinement.}
\def\crvf{R45 May cause cancer.}
\def\crvs{R46 May cause heritable genetic damage.}
\def\crvi{R47 May cause birth defects.}
\def\crva{R48 Danger of serious damage to health by prolonged
exposure.}
\def\crvn{R49 May cause cancer by inhalation.}
\def\crfo{R50 Very toxic to aquatic organisms.}
\def\crfe{R51 Toxic to aquatic organisms.}
\def\crfz{R52 Harmful to aquatic organisms.}
\def\crfd{R53 May cause long-term adverse effects in the aquatic
environment.}
\def\crfv{R54 Toxic to flora.}
\def\crff{R55 Toxic to fauna.}
\def\crfs{R56 Toxic to soil organisms.}
\def\crfi{R57 Toxic to bees.}
\def\crfa{R58 May cause long-term adverse effects in the
environment.}
\def\crfn{R59 Dangerous for the ozone layer.}

% ***********Don't know the translations to the following**************
% * If anyone knows, please mail me <nz@thiemo.net>   *
% \def\crso{R60 Kann die Fortpflanzungsf"ahigkeit beeintr"achtigen.}
% \def\crse{R61 Kann das Kind im Mutterleib sch"adigen.}
% \def\crsz{R62 Kann m"oglicherweise die Fortpflanzungsf"ahigkeit
% beeintr"achtigen}
% \def\crsd{R63 Kann das Kind im Mutterleib m"oglicherweise sch"adigen.}
% \def\crsv{R64 Kann S"auglinge "uber die Muttermilch sch"adigen.}
% **********************************************************************

% Combinations of R-phrases
\def\crcevef{R14/15 Reacts violently with water, liberating highly
flammable gases.}
\def\crcefzn{R15/29 Contact with water liberates toxic, highly
flammable gas.}
\def\crczoze{R20/21 Harmful by inhalation and in contact with skin.}
\def\crczozz{R20/22 Harmful by inhalation and if swallowed.}
\def\crczozezz{R20/21/22 Harmful by inhalation, in contact with
skin and if swallowed.}
\def\crczdzv{R23/24 Toxic by inhalation and in contact with skin.}
\def\crczvzf{R24/25 Toxic in contact with skin and if swallowed.}
\def\crczdzf{R23/25 Toxic by inhalation and if swallowed.}
\def\crczdzvzf{R23/24/25 Toxic by inhalation, in contact with skin
and if swallowed.}
\def\crczszi{R26/27 Very toxic by inhalation and in contact with
skin.}
\def\crcziza{R27/28 Very toxic in contact with skin and if
swallowed.}
\def\crczsza{R26/28 Very toxic by inhalation and if swallowed.}
\def\crczsziza{R26/27/28 Very toxic by inhalation, in contact with
skin and if swallowed.}
\def\crcdsdi{R36/37 Irritating to eyes and respiratory system.}
\def\crcdida{R37/38 Irritating to respiratory system and skin.}
\def\crcdsda{R36/38 Irritating to eyes and skin.}
\def\crcdsdida{R36/37/38 Irritating to eyes, respiratory system and
skin.}
\def\crcvzvd{R42/43 May cause sensitization by inhalation and skin
contact.}
%
% also some 'S-phrases'
% S1-9
\def\cse{S1 Keep locked up.}
\def\csz{S2 Keep out of reach of children.}
% unfortunately, the following is not 'Keep cool' ;-)
\def\csd{S3 Keep in a cool place.}
\def\csv{S4 Keep away from living quarters.}
\def\csf#1{S5 Keep contents under #1 .}
\def\css#1{S6 Keep contents under  #1 .}
\def\csi{S7 Keep container tightly closed.}
\def\csa{S8 Keep container dry.}
\def\csn{S9 Keep container in a well-ventilated place.}
% S10-19
\def\cseo{S10 Keep contents wet.}
\def\csee{S11 Avoid contact with air.}
\def\csez{S12 Do not keep the container sealed.}
\def\csed{S13 Keep away from food, drink and animal feeding stuffs.}
\def\csev#1{S14 Keep away from #1 .}
\def\csef{S15 Keep away from heat.}
\def\cses{S16 Keep away from sources of ignition - No smoking.}
\def\csei{S17 Keep away from combustible material.}
\def\csea{S18 Handle and open container with care.}
% S20-29
\def\cszo{S20 When using do not eat or drink.}
\def\csze{S21 When using do not somoke.}
\def\cszz{S22 Do not breathe dust.}
\def\cszd{S23 Do not breathe gas/fumes/vapour/spray.}
\def\cszv{S24 Avoid contact with skin.}
\def\cszf{S25 Avoid contact with eyes.}
\def\cszs{S26 In case of contact with eyes rinse immediately with
plenty of water and seek medical advice.}
\def\cszi{S27 Take off immediately all contaminated clothing.}
\def\csza#1{S28 After contact with skin wash immediately with
plenty of #1 .}
\def\cszn{S29 Do not empty into drains.}
% S30-39
\def\csdo{S30 Never add water to this product.}
\def\csde{S31 Keep away from explosive materials.}
\def\csdd{S33 Take precautionary measures against static discharges.}
\def\csdv{S34 Avoid shock and friction.}
\def\csdf{S35 This material and its container must be disposed of
in a safe way.}
\def\csds{S36 Wear suitable protective clothing.}
\def\csdi{S37 Wear suitable gloves.}
\def\csda{S38 In case of insufficient ventilation wear suitable
respiratory equipment.}
\def\csdn{S39 Wear eye/face protection.}
% S40-49
\def\csvo#1{S40 To clean the floor and all objects contaminated by
this material use  #1 .}
\def\csve{S41 In case of fire and/or explosion do not breathe fumes.}
\def\csvz{S42 During fumigation/spraying wear suitable respiratory
equipment.}
\def\csvd#1{S43 In case of fire use #1 .}
\def\csvv{S44 If you feel unwell seek medical advice (show the
label where possible).}
\def\csvf{S45 In case of accident or if you feel unwell seek
medical advice immediately (show the label where possible).}
\def\csvs{S46 If swallowed seek medical advice immediately and show
this container or label.}
\def\csvi#1{S47 Keep at temperature not exceeding #1$^\mathrm{o}$C.}
\def\csva#1{S48 Keep wetted with #1 .}
\def\csvn{S49 Keep only in the original container.}
% S50-59
\def\csfo#1{S50 Do not mix with #1 .}
\def\csfe{S51 Use only in well-ventilated areas.}
\def\csfz{S52 Not recommended for interior use on large surface
areas.}
\def\csfd{S53 Avoid exposure - obtain special instructions before
use.}
\def\csfv{S54 Obtain the consent of pollution control authorities before
discharging to wastewater treatment plants.}
\def\csff{S55 Treat using the best available techniques before discharge
into drains or the aquatic environment.}
\def\csfs{S56 Do not discharge into drains or the environment;
dispose to an authorised waste collection point.}
\def\csfi{S57 Use appropriate containment to avoid environmental
contamination.}
\def\csfa{S58 To be disposed of as hazardous waste.}
\def\csfn{S59 Refer to manufacturer/supplier for information on
recovery/recycling.}
\def\csso{S60 This material and/or its container must be disposed
of as hazardous waste.}

% Kombinationen der S-Saetze.
\def\cscez{S1/2 Keep locked up and out of reach of children.}
\def\cscsa{S7/8 Keep container tightly closed and dry.}
\def\cscdsn{S3/7/9 Keep container tightly closed in a cool,
well-ventilated place.}
\def\cscdn{S3/9 Keep in a cool, well-ventilated place.}
\def\cscdev#1{S3/14 Keep in a cool place away from #1 .}
\def\cscdnev#1{S3/9/14 Keep in a cool, well-ventilated place away
from #1 .}
\def\cscdnvn{S3/9/49 Keep only in the original container in a cool,
well-ventilated place.}
\def\cscdnevvn#1{S3/9/14/49 Keep only in the original container in
a cool, well-ventilated place away from #1 .}
\def\cscia{S7/8 Keep container tightly closed and dry.}
\def\cscin{S7/9 Keep container tightly closed and in a
well-ventilated place.}
\def\csczoze{S20/21 When using do not eat drink or smoke.}
\def\csczvzf{S24/25 Avoid contact with skin and eyes.}
\def\cscdsdi{S36/37 Wear suitable protective clothing and gloves.}
\def\cscdsdn{S36/39 Wear suitable protective clothing and eye/face
protection.}
\def\cscdidn{S37/39 Wear suitable gloves and eye/face protection.}
\def\cscdsdidn{S37/38/39 Wear suitable gloves, respiratory
equipment and eye/face protection.}
\def\cscvivn#1{S47/49 Keep only in the original container at
temperature not exceeding #1$^\mathrm{o}$C.}

% -- Pooh! got it :)
